---
title: "STAT 33A Workbook 5"
date: "Oct 1, 2020"
author: "CJ HINES (3034590053)"
output: pdf_document
---

This workbook is due __Oct 1, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.

If you have any last-minute trouble knitting, **DON'T PANIC**. Submit your Rmd
file on time and follow up in office hours or on Piazza to sort out the PDF.


R Graphics Overview
===================

Watch the "R Graphics Overview" lecture video.

No exercises for this section. Keep at it!


The Tidyverse
=============

Watch the "The Tidyverse" lecture video.


## Exercise 1

How many packages are in the Tidyverse? Explore the website to find out. You
can count the tidymodels packages as a single package.

**YOUR ANSWER GOES HERE:**

> There are 22 packages in Tidyverse.


Tibbles
=======

Watch the "Tibbles" lecture video.

## Exercise 2

1. Read the documentation for the tibble package on the website. What's the
   name of the function that creates a new tibble from column vectors?

2. Create a tibble with 4 rows and 3 columns. You can make up the data in the
   columns, but use a different data type for each one.

3. Show how to convert the tibble from step 2 into an ordinary data frame.

**YOUR ANSWER GOES HERE:**

1. The function that creates a new tibble from column vectors is `tibble()`.
```{r}
library(tibble)
tibble()
```

2.
```{r}
atib = tibble(x = 1:4, y = c(3.3, 4.0, 5.0, 6.2), z = "canada")
atib
```

3.
```{r}
class(atib)
class(as.data.frame(atib))
```



The Grammar of Graphics
=======================

Watch the "The Grammar of Graphics" lecture video.

For exercises that mention the dogs data, you can use either `dogs.rds` or
`dogs_tibble.rds` from the bCourse (do not use `dogs_sample.rds`).


## Exercise 3

Use ggplot2 and the dogs data to make a scatterplot that shows the relationship
between height and weight. In 2-3 sentences, discuss any patterns you see in
the plot.

**YOUR ANSWER GOES HERE:**

```{r}
library(ggplot2)
dogs = readRDS("dogs.rds")

ggplot(dogs, aes(x = height, y = weight)) + geom_point()
```

> A pattern from the plot is that as height increases, weight increases. It's an almost linear graph.

## Exercise 4

The `geom_bar()` geometry function makes a bar plot. Even though a bar plot has
two axes, by default `geom_bar()` only requires one aesthetic: `x`. The y-axis
is computed automatically and shows the frequencies of the values on the
x-axis. This makes `geom_bar()` especially convenient for displaying
categorical data.

1. Make a bar plot that shows the number of dogs in each "group" of dogs.

2. Are any groups much larger or smaller than the others?

**YOUR ANSWER GOES HERE:**

1. 
```{r}
ggplot(dogs, aes(x = group)) + geom_bar()
```

2. The sporting,terrier, and working groups are the largest groups. The non-sporting and toy groups are the smallest.


## Exercise 5

1. Which geometry function makes a histogram? Use the ggplot2 website or cheat
   sheet to find out.

2. Use ggplot2 to make a histogram of longevity for the dogs data. How long do
   most dogs typically live? How spread out is the distribution of lifespans?

   _Hint 1: If you're not familiar with histograms, OpenIntro Statistics 4th
   Edition (link on Piazza) explains how to interpret them in Section 2.1.3._

   _Hint 2: Describe how "spread out" the distribution is at whatever level of
   statistics you're comfortable with. Range is the simplest, but standard
   deviation or other statistics are also fine._

**YOUR ANSWER GOES HERE:**

1. `geom_histogram()`

2. Most dogs typically live around 9-12 years. The spread out of the distribution of lifespans is 10.21.
```{r}
ggplot(dogs, aes(x = longevity)) + geom_histogram()
range(dogs$longevity, na.rm = TRUE)
#sort(dogs$longevity)
```


Saving Plots
============

Watch the "Saving Plots" lecture video.

No exercises for this section. Almost done!


Customizing Plots
=================

Watch the "Customizing Plots" lecture video.


## Exercise 6

1. Modify your plot from Exercise 3 so that the shape of the points is
   determined by the "group" of the dog.

   _Hint: To do this, you need to set another aesthetic. The documentation for
   `geom_point()` includes an example that shows how to change the shape of the
   points._

**YOUR ANSWER GOES HERE:**

1. 
```{r}
ggplot(dogs, aes(x = height, y = weight, shape = factor(group))) + geom_point() + scale_shape_manual(values=seq(0, 7-1))
```


## Exercise 7

1. Modify your plot from Exercise 6 so that the color of the points is *also*
   determined by the "group" of the dog.

2. Features that separate different categories of observations are useful for
   building models that predict those categories.

   Do height and weight effectively separate the different groups of dogs? In
   other words, are there clear boundaries between the groups in the plot (as
   opposed to being mixed together)? Are some groups better separated than
   others?

**YOUR ANSWER GOES HERE:**

1. 
```{r}
ggplot(dogs, aes(x = height, y = weight, shape = factor(group))) + geom_point(aes(color = factor(group))) + scale_shape_manual(values=seq(0, 7-1))
```

2. Height and weight separate some groups more effectively than others. For example, dogs of the working group are mostly very heavy and tall and dogs of the toy group are mostly short and light. For dog groups like hound and non-sporting, though, there is a bit more uncertainty and mix.
