---
title: "STAT 33A Workbook 7"
date: "Oct 15, 2020"
author: "CJ HINES (3034590053)"
output: pdf_document
---

This workbook is due __Oct 15, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.

If you have any last-minute trouble knitting, **DON'T PANIC**. Submit your Rmd
file on time and follow up in office hours or on Piazza to sort out the PDF.


dplyr Overview
==============

Watch the "dplyr Overview" lecture video.

No exercises for this section.


Subsets with dplyr
==================

Watch the "Subsets with dplyr" lecture video.

## Exercise 1

Use dplyr and the dogs data to compute each of the following subsets:

1. Rows 10-30 only

2. All rows except row 51

3. All columns except `popularity_all` and `popularity`

4. Rows 1-10 with only the `breed`, `weight`, and `height` columns

You do not need to print out these subsets, just show us the code to compute
them.

**YOUR ANSWER GOES HERE:**
```{r}
library(dplyr)
dogs = readRDS("dogs.rds")
```

1. Rows 10-30 only
```{r}
#slice(dogs, seq(10,30))
tentothirty = slice(dogs, 10:30)
```

2. All rows except row 51
```{r}
no51 = slice(dogs, -51)
```

3. All columns except `popularity_all` and `popularity`
```{r}
nopop = select(dogs, -popularity_all, -popularity)
```

4. Rows 1-10 with only the `breed`, `weight`, and `height` columns
```{r}
dogsbwh = select(dogs, breed, weight, height)
sldogsbwh = slice(dogsbwh, 1:10)
```

## Exercise 2

Use dplyr to show that there are no duplicated rows in the dogs data.

Explain your reasoning.

**YOUR ANSWER GOES HERE:**
```{r}
howmany = nrow(dogs)
distincthowmany = nrow(distinct(dogs))
howmany == distincthowmany
```

`howmany` are the number of rows in dogs. `distincthowmany` are the number of distinct (or not duplicated) rows in dogs. The numbers are equal.


Base R versus dplyr
===================

Watch the "Base R versus dplyr" lecture video.

No exercises for this section.


Transformations with dplyr
==========================

Watch the "Transformations with dplyr" lecture video.


## Exercise 3

Workbook 4, Exericse 6 asked you to use base R and the dogs data to compute:

1. The mean and median of the longevity column (ignoring missing values).

2. The subset that contains rows 10-20 of the height, weight, and longevity
   columns.

3. The number of dog breeds with weight greater than 42.

4. The subset of large dogs that require daily grooming.

For each of these, show the code to compute the result:

1. Using base R
2. Using dplyr


**YOUR ANSWER GOES HERE:**

**BASE R**

1. The mean and median of the longevity column (ignoring missing values).
```{r}
mean(dogs$longevity, na.rm = TRUE)
median(dogs$longevity, na.rm = TRUE)
```

2. The subset that contains rows 10-20 of the height, weight, and longevity columns.
```{r}
rh = dogs$height[10:20]
rw = dogs$weight[10:20]
rl = dogs$longevity[10:20]
```

3. The number of dog breeds with weight greater than 42.
```{r}
heavy = subset(dogs, weight > 42)
length(heavy$breed)
```

4. The subset of large dogs that require daily grooming.
```{r}
biggroom = subset(dogs, size == "large" & grooming == "daily")
```

**DPLYR**

1. The mean and median of the longevity column (ignoring missing values).
```{r}
summarize(dogs, mean(longevity, na.rm = TRUE))
summarize(dogs, median(longevity, na.rm = TRUE))
```

2. The subset that contains rows 10-20 of the height, weight, and longevity columns.
```{r}
tentwenty = slice(dogs, 10:20)
select(tentwenty, height, weight, longevity)
```

3. The number of dog breeds with weight greater than 42.
```{r}
count(filter(dogs, weight > 42))
```

4. The subset of large dogs that require daily grooming.
```{r}
filt = filter(dogs, size == "large", grooming == "daily")
```


## Exercise 4

Use dplyr and the dogs data to determine which 3 dogs cost the most.

Your answer to this exercise should be a data frame with 3 rows.

**YOUR ANSWER GOES HERE:**
```{r}
dogs = readRDS("dogs.rds")
filter(dogs, lifetime_cost > 25600)
```


## Exercise 5

Use dplyr to answer each of the following:

1. On average, which `group` of dog has the highest lifetime cost? Which has
   the lowest?

2. How many dogs are there for each possible combination of `size` and
   `grooming`?

3. For each `group` of dog, what's the shortest lifespan? You should have one
   result per group here. For each `group` of dog, what's the longest lifespan?

4. Do popular dogs tend to be more expensive? Use any columns that seem
   appropriate; you can also use ggplot2 if you like.


**YOUR ANSWER GOES HERE:**

1. On average, which `group` of dog has the highest lifetime cost? Which has the lowest?
On average, herding dogs have the highgest lifetime costs while working dogs have the lowest.
```{r}
bygroup = group_by(dogs, group)
summarize(bygroup, mean(lifetime_cost, na.rm = TRUE))
```

2. How many dogs are there for each possible combination of `size` and `grooming`? 
```{r}
count(dogs, size, grooming)
```

3. For each `group` of dog, what's the shortest lifespan? You should have one
   result per group here. For each `group` of dog, what's the longest lifespan?

Shortest:
```{r}
summarize(bygroup, min(longevity, na.rm = TRUE))
```

Longest:
```{r}
summarize(bygroup, max(longevity, na.rm = TRUE))
```

4. Do popular dogs tend to be more expensive? Use any columns that seem
   appropriate; you can also use ggplot2 if you like.
   
> No, popularity is not related to lifetime_cost.
   
```{r}
library(ggplot2)
ggplot(dogs, aes(popularity_all, lifetime_cost, color = group)) + geom_point()
```


The Pipe Operator
=================

Watch the "The Pipe Operator" lecture video.

## Exercise 6

Read the documentation for the pipe operator at
<https://magrittr.tidyverse.org/reference/pipe.html>.

How do you pass the left-hand operand as the *second* argument to the
right-hand side?

Give an example of computing the logarithm of 3, base 10, where you use the
pipe to pass `10` as the argument for `base`.

**YOUR ANSWER GOES HERE:**
```{r}
library("dplyr")
3 %>% log(10)
```
