---
title: "STAT 33A Workbook 3"
date: "Sep 17, 2020"
author: "CJ Hines (3034590053)"
output: pdf_document
---

This workbook is due __Sep 17, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.


File Systems
============

Watch the "File Systems" lecture video.


## Exercise 1

For each of the following paths, say whether the path is absolute or relative,
and explain how you can tell.

1. `"/Users/Jun/doggos_to_pet.md"`
2. `"TODO.md"`
3. `"./documents"`

YOUR ANSWER GOES HERE:

> 1. The path is absolute, it starts from the root. 
> 2. This path is relative, it does not begin at the root.
> 3. This path is relative,  it starts from the current directory.


The R Working Directory
=======================

Watch the "The R Working Directory" lecture video.


## Exercise 2

1. What's the root directory called on your computer?
2. What's the absolute path to the home directory on your computer?
3. Are the root directory and the home directory different directories? How can
   you tell?
4. Use R to count the total number of files and directories in your home
   directory. Your code should return the result as a number. _Hint: the
   `length()` function returns the length of a vector._

YOUR ANSWER GOES HERE:

1. `/`
2. `/Users/cj`
3. They are different because `/` is just the beginning of the path to `/Users/cj`.
4. 16

```{r}
x = list.files("/Users/cj")
length(x)
```

Data Frames
===========

Watch the "Data Frames" lecture video.

## Exercise 3

1. Load the dogs data set `dogs.rds` into R (this one has more than 10 rows).
2. What's the mean weight of the dogs? You can use the `na.rm` parameter in the
   `mean()` function to make the function ignore missing values.
3. The `which.min()` function returns the index of the minimum element of a
   vector. Find the row number for the dog with the shortest height, then use
   subsetting to get the breed of that dog.
4. Which breed of dog has the longest lifespan? _Hint: min/max functions come
   in pairs._

YOUR ANSWER GOES HERE:

1. 
```{r}
dogs = readRDS("dogs.rds")
```
2.
```{r}
mean(dogs$weight, na.rm = TRUE)
```
3. 
```{r}
shortestrow = which.min(dogs$height)
dogs$breed[shortestrow]
```
4. 
```{r}
long = which.max(dogs$longevity)
dogs$breed[long]
```


Factors
=======

Watch the "Factors" lecture video.


No exercises for this video. Get up, stretch, and take a break! :)


File Formats
============

Watch the "File Formats" lecture video.


## Exercise 4

1. Load the volcano data set into R.
2. What are the column names? Use R to get these rather than typing them out
   yourself.
3. How many volcano eruptions are recorded in the data set?
4. What are the classes of the columns? _Hint: the "Data Frames" lecture
   included a function that summarizes of this information._
5. Statistically, which of the columns in the data set are categorical?
6. In R, which of the columns in the data set are categorical? How can you
   tell?
7. How do your answers in part 5 and part 6 compare? Does R correctly infer
   which columns are categorical?

YOUR ANSWER GOES HERE:

1. 
```{r}
volcs = read.delim("volerup.txt")
```
2. 
```{r}
columnnames = names(volcs)
columnnames
```
3. 
```{r}
howmany = length(rownames(volcs))
howmany
```
4. int, num, chr
```{r}
str(volcs)
```
5. Name, Location, Country, Type, Status, Time, Agent.
```{r}
str(volcs)
```
6. TSU,  EQ,  Name, Location, Country, Type, Status, Time, Agent Their class is chr.
```{r}
str(volcs)
```
7. They're about the same. Yes.
```{r}
str(volcs)
```